---
title: "Assignment 1"
author: "Kristy Henrich"
date: "9/6/2020"
output: html_document
---

``` {r load libraries, message = FALSE, results = 'hide'}
library(tidycensus)
library(tidyverse)
```

## Exploring which variables to use 

I explored all three levels of data tables from the 2018 ACS 5-Year estimates to understand the variables available. I narrowed down to the tract-level analysis from the ACS 5-Year estimates because I wanted to include estimates for geographic areas with populations of fewer than 20,000 people, a tip I learned from Cat's assignment. 

```{r load person-level microdata variables}
person_vars_2018 <- pums_variables %>%
  distinct(year, survey, var_code, 
           var_label, data_type, level) %>%
  filter(level == "person", year == 2018, survey == "acs5")
```

```{r load household-level variables}
hh_vars_2018 <- pums_variables %>%
  distinct(year, survey, var_code, 
           var_label, data_type, level) %>%
  filter(level == "housing", year == 2018, survey == "acs5")
```

``` {r load area-level variables}
area_vars_2018 <- load_variables(2018, "acs5")
```

## Loading the data sets

I decided to conduct a census-tract level analysis in the nine-county SF Bay Area to explore internet access trends. In my exploration, I would also like to understand how internet access differs or not by the county.

First, I had to load all of the variables that I would need for my data set and any calculations to create my data set. After loading my variables, I created a data set for each individual county of the nine-county Bay Area so that I could add a column to my data set to denote county. Finally, I added the county column to each individual county data set.

Unit of analysis: census-tract

  1. Bay Area county (categorical) : county
  2. Majority race / ethnicity of the population (categorical) : maj_re
  3. Whether the majority of the units are occupied by 
      renters vs. owners (categorical) : unit_type
  4. Percent of population that has no internet access 
      (continuous) : no_internet
  5. Median income (continuous) : med_income
  6. Average household size (continuous): avg_hh_size


``` {r load 2018 ACS-5 Year variables for analysis, message=FALSE, include = TRUE, results = 'hide'}
pop_vars <- c(total_pop =         "B02001_001", 
              white_alone =       "B03002_003",
              black_AA_alone =    "B03002_004",
              native_alone =      "B03002_005",
              asian_alone =       "B03002_006",
              hawaiian_PI_alone = "B03002_007",
              other =             "B03002_008",
              mult_races =        "B03002_009",
              hispanic_latino =   "B03002_012",
              
              total_units =       "B25003_001",
              units_rented =      "B25003_003",
              units_owned =       "B25003_002",
              
              tot_hh_internet =   "B28011_001",
              no_internet =       "B28011_008",
              
              med_income =        "B06011_001",
              
              avg_hh_size =       "B25010_001")

tractAlameda2018 <- get_acs(geography= "tract", 
              variables = pop_vars, 
              output = "wide", 
              state = "CA",
              county = "Alameda",
              year = 2018, 
              survey = "acs5")

tractAlameda2018$county <- rep("Alameda", times = length(tractAlameda2018$total_popE))

tractContraCosta2018 <- get_acs(geography= "tract", 
              variables = pop_vars, 
              output = "wide", 
              state = "CA",
              county = "Contra Costa",
              year = 2018, 
              survey = "acs5")

tractContraCosta2018$county <- rep("ContraCosta", times = length(tractContraCosta2018$total_popE))

tractMarin2018 <- get_acs(geography= "tract", 
              variables = pop_vars, 
              output = "wide", 
              state = "CA",
              county = "Marin",
              year = 2018, 
              survey = "acs5")

tractMarin2018$county <- rep("Marin", times = length(tractMarin2018$total_popE))

tractNapa2018 <- get_acs(geography= "tract", 
              variables = pop_vars, 
              output = "wide", 
              state = "CA",
              county = "Napa",
              year = 2018, 
              survey = "acs5")

tractNapa2018$county <- rep("Napa", times = length(tractNapa2018$total_popE))

tractSanFrancisco2018 <- get_acs(geography= "tract", 
              variables = pop_vars, 
              output = "wide", 
              state = "CA",
              county = "San Francisco",
              year = 2018, 
              survey = "acs5")

tractSanFrancisco2018$county <- rep("San Francisco", times = length(tractSanFrancisco2018$total_popE))

tractSanMateo2018 <- get_acs(geography= "tract", 
              variables = pop_vars, 
              output = "wide", 
              state = "CA",
              county = "San Mateo",
              year = 2018, 
              survey = "acs5")

tractSanMateo2018$county <- rep("San Mateo", times = length(tractSanMateo2018$total_popE))

tractSantaClara2018 <- get_acs(geography= "tract", 
              variables = pop_vars, 
              output = "wide", 
              state = "CA",
              county = "Santa Clara",
              year = 2018, 
              survey = "acs5")

tractSantaClara2018$county <- rep("Santa Clara", times = length(tractSantaClara2018$total_popE))

tractSolano2018 <- get_acs(geography= "tract", 
              variables = pop_vars, 
              output = "wide", 
              state = "CA",
              county = "Solano",
              year = 2018, 
              survey = "acs5")

tractSolano2018$county <- rep("Solano", times = length(tractSolano2018$total_popE))

tractSonoma2018 <- get_acs(geography= "tract", 
              variables = pop_vars, 
              output = "wide", 
              state = "CA",
              county = "Sonoma",
              year = 2018, 
              survey = "acs5")

tractSonoma2018$county <- rep("Sonoma", times = length(tractSonoma2018$total_popE))

```

## Joining county-level datasets
Once I had the data from each individual county, I merged it into one data set.

```{r}
bayareaCA2018 <- rbind(tractAlameda2018, tractContraCosta2018, tractMarin2018, tractNapa2018, tractSanFrancisco2018, tractSanMateo2018, tractSantaClara2018, tractSolano2018, tractSonoma2018)
```

## Calculating new variables
Once I brought in all variables that I need for calculations, I calculated the variables I would need in my analysis (majority race/ethnicity, whether the majority of units were rented, and percent of households with no internet access). I used Cat's code to convert race/ethnicity and mirrored her logic for my other calculations.

```{r}
bayareaCA2018 <- bayareaCA2018 %>%
  mutate(pct_white = white_aloneE/total_popE,
         pct_black = black_AA_aloneE/total_popE,
         pct_native = native_aloneE/total_popE,
         pct_asian = asian_aloneE/total_popE,
         pct_hawaiian_PI = hawaiian_PI_aloneE/total_popE,
         pct_mult = mult_racesE/total_popE,
         pct_other = otherE/total_popE,
         pct_hisp_lat = hispanic_latinoE/total_popE,
         maj_units_rented = units_rentedE/total_unitsE > 0.5,
         pct_no_internet = no_internetE/tot_hh_internetE)%>%
         
 mutate(maj_re = case_when(pct_white > 0.5 ~ "nh_white",
                             pct_black > 0.5 ~ "nh_black",
                             pct_native > 0.5 ~ "native_am",
                             pct_asian > 0.5 ~ "nh_asian",
                             pct_hawaiian_PI > 0.5 ~ "hawaiian_PI",
                             pct_mult > 0.5 ~ "multi",
                             pct_other > 0.5 ~ "other_nh",
                             pct_hisp_lat > 0.5 ~ "hisp_latino",
                             TRUE ~ "no_maj_re")) %>%

select (county, maj_re, maj_units_rented, pct_no_internet, 
          med_incomeE, avg_hh_sizeE)
```

## Loading the data table
Once I had the variables calculated and final ones for the table selected, I loaded the data table to test the results.

```{r}
bayareaCA2018
```

## Showing the first 10 rows
Finally, I pulled up the first 10 rows for the assignment. To recap, I am conducting an analysis on the following:

  1. unit of analysis : tract-level
  2. definition of sample population : census tracts in the nine-county Bay Area
  3. variables in analysis : Bay Area county (categorical), majority race / ethnicity of the population (categorical), whether the majority of the units are occupied by renters vs. owners (categorical), percent of population that has no internet access (continuous), median income (continuous), average household size (continuous)
  4. number of observations in data set: 1,588 observations of 6 variables

```{r}
head(bayareaCA2018, 10)
```
```{r}
write.csv(bayareaCA2018)
```
